Topic categorization in Portuguese news articles\\ \thanks{Fundação para a Ciência e Tecnologia}

### 2018-01-03 article.tex
Contributions and Future Work

%TODO
A document-term matrix was then calculated using the
\texttt{DocumentTermMatrix} method from the \texttt{tm} package.
For each document in the train set the matrix included all the
terms which met the following requirements:

@@

The test dataset was preprocessed in a way similar to the train set
(described in the previous section,~\ref{sec:preproc}),
but the final list of terms in each test document was restricted to the
terms considered for the train set.

\section{Results}

Table~\ref{tab:size} presents the size of each list of terms after
removing the

\begin{tabular}{S[table-format=3.0]}
\toprule
{Heading} \\
\midrule
1 \\
10 \\
100 \\
\bottomrule
\end{tabular}

@@

@@

\begin{table}[htbp]
    \caption{Execution times for training}
\begin{center}
\begin{tabular}{lcr}
                 & Algorithm  & Execution time \\\hline
    $LT_{90}$    & DT         & 12s  \\
    $LT_{95}$    & KNN        &  3m  \\
    $LT_{99}$    & KNN        & 57m  \\
\end{tabular}
\label{tab:exectime}
\end{center}
\end{table}

@@

\begin{table}[htbp]
    \caption{Categories confusion matrix ($LT_{99}$ with KNN)}
\begin{center}
\begin{tabular}{l|rrrrrrrr}
& \rotatebox{75}{ciencia}    & \rotatebox{75}{cultura}  &
    \rotatebox{75}{desporto} & \rotatebox{75}{economia} &
    \rotatebox{75}{mundo}    & \rotatebox{75}{musica}   &
    \rotatebox{75}{pais}     & \rotatebox{75}{politica} \\\hline
ciencia  & {\bf 127} & 19        & 1          & 6         & 12       & 24        & 11       & 0 \\
cultura  & 5         & {\bf 102} & 2          & 2         & 3        & 79        & 6        & 1 \\
desporto & 1         & 3         & {\bf  168} & 7         & 3        & 14        & 4        & 0 \\
economia & 5         & 3         & 0          & {\bf 146} & 5        & 17        & 14       & 10 \\
mundo    & 12        & 12        & 7          & 17        & {\bf 90} & 35        & 16       & 11 \\
musica   & 0         & 2         & 0          & 0         & 2        & {\bf 184} & 1        & 0 \\
pais     & 9         & 9         & 9          & 19        & 15       & 31        & {\bf 88} & 21 \\
politica & 1         & 3         & 0          & 30        & 8        & 17        & 14       & {\bf 127} \\
\end{tabular}
\label{tab:confmat}
\end{center}
\end{table}

\section{Discussion}
% nested categories
% or not mutex categories

Looking to the error rate obtained in each of tests performed, we can
conclude that the best option was to use the largest list of terms
($LT_{99}$) and the k-nearest neighbors classifier.

>>>>

%TODO
A document-term matrix was then calculated using the
\texttt{DocumentTermMatrix} method from the \texttt{tm} package.
For each of the 5600 documents preset in the train set the matrix
included all the terms which met the following requirements:

@@

The test dataset contained 1600 documents (200 belonging to each
category) and was preprocessed in a way similar to the train set
(described in the previous section,~\ref{sec:preproc}).  In each
iteration, however, the final list of terms in each test document was
restricted to terms present in the corresponding list of terms
($LT_{99}$, $LT_{95}$ and $LT_{90}$).

\section{Results}
A number of measurements and metrics were calculated regarding the
lists of terms, the classifiers training process and their results in
the evaluation process.

Table~\ref{tab:size} presents the size of each list of terms after
removing the terms whose sparsity was above the corresponding
threshold.

@@

\begin{table}[htbp]
    \caption{Execution times for training}
\begin{center}
\begin{tabular}{l|cr}
    & Algorithm  & \parbox{3em}{\centering Training\\time} \\\hline
    $LT_{90}$    & DT         & 12s  \\
    $LT_{95}$    & KNN        &  3m  \\
    $LT_{99}$    & KNN        & 57m  \\
\end{tabular}
\label{tab:exectime}
\end{center}
\end{table}

@@

@@

\begin{table}[htbp]
    \caption{Categories confusion matrix ($LT_{99}$ with KNN)}
\begin{center}
\begin{tabular}{l|rrrrrrrr}
& \rotatebox{75}{science}    & \rotatebox{75}{culture}  &
    \rotatebox{75}{sports} & \rotatebox{75}{economics} &
    \rotatebox{75}{international}    & \rotatebox{75}{music}   &
    \rotatebox{75}{national}     & \rotatebox{75}{politics} \\\hline
science  & {\bf 127} & 19        & 1          & 6         & 12       & 24        & 11       & 0 \\
culture  & 5         & {\bf 102} & 2          & 2         & 3        & 79        & 6        & 1 \\
sports   & 1         & 3         & {\bf  168} & 7         & 3        & 14        & 4        & 0 \\
economics & 5         & 3         & 0          & {\bf 146} & 5        & 17        & 14       & 10 \\
international    & 12        & 12        & 7          & 17        & {\bf 90} & 35        & 16       & 11 \\
music   & 0         & 2         & 0          & 0         & 2        & {\bf 184} & 1        & 0 \\
national     & 9         & 9         & 9          & 19        & 15       & 31        & {\bf 88} & 21 \\
politics & 1         & 3         & 0          & 30        & 8        & 17        & 14       & {\bf 127} \\
\end{tabular}
\label{tab:confmat}
\end{center}
\end{table}

\section{Discussion}

The analysis of the results obtained should take into account other
metrics besides the error rates obtained for each classifier.

The size of the list of terms, for example, gives us an idea of the
memory footprint of a classifier, a parameter which is of the utmost
importance if the goal is to implement a classifier as lightweight as
possible. The training time is also relevant, as shorter training
times give the possibility of retraining the classifiers more often,
allowing them to be updated as the corpus of news articles grows.

Looking at the actual results obtained and represented in
Tables~\ref{tab:size} and \ref{tab:exectime} we can see that $LT_{90}$
has simultaneously the smallest list of terms (45) and the shortest
training time (12 seconds using the decision tree algorithm). However,
$LT_{90}$ also presents the worst error rates, even when looking at
the algorithm which achieved its best results (0.60 using a decision
tree classifier).

On the opposite side, $LT_{99}$ presented the lowest error rates (0.35
using a k-nearest neighbors classifier), but it took almost one hour
to train and used a list comprising 912 terms.

The results obtained confirm that there is a tradeoff between the size
of the list of terms and the training time, on the one hand, and the
classifier error rate, on the other. However, a list of 912 terms
seems to be an acceptable memory footprint for a client side
classifier; additionally, the additional training time would not present
much problems in this scenario, as the classifier would be trained
before hand and thus would not be visible on the client side.

Given the reasonable values for the training time and the size of the
list of terms, and looking to the error rate obtained in each of tests
performed, we can conclude that the best option was to use the largest
list of terms ($LT_{99}$) and the k-nearest neighbors classifier.

It is worth noting that the categories of an article are not mutually
exclusive. In fact, an article can be classified as belonging to more
than one category. This might explain the greater error rates obtained
in the categories \textit{national} and  \textit{international}: one
might argue that these categories correspond less to the topic covered
by the article and are more related to the location of the news
content.

\section{Contributions and Future work}
For copyright reasons, the corpus used to train and evaluate the
classifiers described in this article cannot be shared. All the code
used to process the documents, to implement the classifiers and
evaluate them can be found at\\
\url{http://github.com/andrefs/mapi-msr-categorization}.

The tasks and results previously described already provide useful
insights into this matter. However, the lowest error rate obtained
(0.35), despite already acceptable, might still be improved, either by
trying new algorithms, fine tuning the ones already tested, or by
increasing the train corpus.

We have established that it is in fact possible
to develop a news article classifier which is lightweight enough to be
used (in real time) on the client side of a web application. The
following step will be the actual implementation, probably in the form
of a JavaScript library, of such a classifier.
